Experiences with Planning for Natural Language Generation

### 2008-06-23 trunk/Papers/experiences/experiments.tex
small grammatical changes

The results of this experiment are shown in
Figure~\ref{fig:runtimes-crisp}. The input parameters $(m,n)$ are plotted
(in lexicographic order) on the horizontal axis and the runtime is shown in
milliseconds on the vertical axis. These results reveal a number of
interesting insights. First, FF significantly outperforms
SGPLAN in this domain.\footnote{Experiments with SGPLAN use a
 pre-release version of SGPLAN, kindly provided by Chih-Wei Hsu. The
 release version of SGPLAN 5.2.2 had a bug, causing it to crash on some
 instances.}
Second, FF's runtime is dominated by its initial grounding step, in which
it computes the ground instances of all predicates and actions used in the
planning problem to avoid unnecessary instantiations during search. In
particular, the ratio of grounding time to total runtime is generally above
85\%, and rises to above 99\% at $m=11$, which is still a small universe in
this application.\footnote{The
  ``grounding'' time reported here is what FF reports as ``time spent:
  instantiating action templates''.}

In fact, FF spends so much time on grounding that it is consistently
outperformed by an ad-hoc Java implementation of GraphPlan which only
computes instances of predicates and actions as they are discovered,
while the planning graph is being built -- despite the fact that FF is
consistently much faster as far as pure search time is concerned.
Looking at the difference from a different angle, FF's performance is
much more sensitive to the domain size: if we fix $n=1$, FF takes 60
ms to compute a plan at $m=1$, but 2.4 seconds (for the same plan) at
$m=10$; our GraphPlan implementation takes 30 ms at $m=1$ and still
only 50 ms at $m=10$.  Conversely, GraphPlan's runtime grows much
faster with the plan size (i.e., with growing values of $n$ for a
fixed $m$).  Larger (but still realistically-sized) instances of the
sentence generation problem are still problematic for the planners we
tested.

@@

In the second experiment, we evaluate the performance of the planners
on problems arising in the GIVE domain. We construct a series of test
worlds, similar to the one illustrated in
Figure~\ref{fig:give-minimal}. These worlds consist of a $2n$ by $h$
grid of positions, such that there are buttons at positions $(2i-1,1)$
and $(2i,h)$ for $1 \leq i \leq n$. The player starts in position
$(1,1)$ and must press all buttons in order to successfully complete
the game. The world is generated as a GIVE world description, and then
automatically converted into a planning problem as in
Fig.~\ref{fig:give-planning} by the GIVE software.

@@

In the final experiment, we vary the structure of the GIVE world in order
to judge the effect that universe size has on the planning problem in this
domain.  Starting with the ordinary GIVE world described in Experiment~2,
we add another $w$ by $h$ empty ``junk'' positions to the right of the
minimal world (see Figure~\ref{fig:give-junk}). These new positions are not
actually needed in any plan, but approximate the situation in the actual
GIVE domain, where most grid positions never need to be used. We leave the
initial state and goal untouched. As before, we generate a GIVE world
description and then convert it into a planning problem.

>>>>

The results of this experiment are shown in
Figure~\ref{fig:runtimes-crisp}. The input parameters $(m,n)$ are plotted
in lexicographic order on the horizontal axis and the runtime is shown in
milliseconds on the vertical axis. These results reveal a number of
interesting insights. First, FF significantly outperforms
SGPLAN in this domain.\footnote{Experiments with SGPLAN use a
 pre-release version of SGPLAN, kindly provided by Chih-Wei Hsu. The
 release version of SGPLAN 5.2.2 had a bug, causing it to crash on some
 instances.}
Second, FF's runtime is dominated by its initial grounding step, in which
it computes the ground instances of all predicates and actions used in the
planning problem to avoid unnecessary instantiations during search. In
particular, the ratio of grounding time to total runtime is generally above
85\%, and rises to above 99\% at $m=11$, which is still a small universe in
this application.\footnote{The
  ``grounding'' time reported here is what FF reports as ``time spent:
  instantiating action templates''.}

In fact, the time spent by FF on grounding is such that it is consistently
outperformed by an ad-hoc Java implementation of GraphPlan which only
computes instances of predicates and actions as they are discovered (i.e.,
while the planning graph is being built)---despite the fact that FF is
consistently much faster as far as pure search time is concerned.  Looking
at the difference from another angle, FF's performance is much more
sensitive to the domain size: if we fix $n=1$, FF takes 60 ms to compute a
plan at $m=1$, but 2.4 seconds (for the same plan) at $m=10$; our GraphPlan
implementation takes 30 ms at $m=1$ and still only requires 50 ms at
$m=10$.  Conversely, GraphPlan's runtime grows much faster with the plan
size (i.e., with growing values of $n$ for a fixed $m$).  Larger (but still
realistically-sized) instances of the sentence generation problem are still
problematic for the planners we tested.

@@

In the second experiment, we evaluate the performance of the planners
on problems arising in the GIVE domain. We construct a series of test
worlds, similar to the one illustrated in
Figure~\ref{fig:give-minimal}. These worlds consist of a $2n$ by $h$
grid of positions, such that there are buttons at positions $(2i-1,1)$
and $(2i,h)$ for $1 \leq i \leq n$. The player starts in position $(1,1)$
and must press all the buttons in order to successfully complete the game.
The world is generated as a GIVE world description, and then automatically
converted into a planning problem (as in Figure~\ref{fig:give-planning}) by
the GIVE software.

@@

In the final experiment, we vary the structure of the GIVE world in order
to judge the effect of universe size on the planning problem in this
domain.  Starting with the ordinary GIVE world described in Experiment~2,
we add another $w$ by $h$ empty ``junk'' positions to the right of the
minimal world (see Figure~\ref{fig:give-junk}). These new positions are not
actually needed in any plan, but approximate the situation in the actual
GIVE domain, where most grid positions are never used. We leave the initial
state and goal untouched. As before, we generate a GIVE world description
and then convert it into a planning problem.
